---
title: "AllRecipes 2: Isolating Ingredients"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


This post continues using the AllRecipes recipe parser from the previous post.


Libraries used:


```{r, warning = FALSE, message = F}

# data wrangling
library(tidyverse)  
# string stuff
library(stringr)   
# pipes
library(magrittr)
# plots
library(ggplot2)
```


#Isolating a single ingredient across recipes

First we try to isolate a single ingredient, here "brown sugar", so that we might compare the related (chocolate chip cookie) recipes later:

```{r}

#initialize a vector the length of the ingredients list-of-ingredients lists to store a single ingredient
brown_sugar <- vector(mode="character", length=length(ingredients))

#loop over the list-of-ingredients lists
for (i in 1:length(ingredients)){
  #on each individual ingredients 
  z <- ingredients[[i]]
  #isolate the single ingredient of interest
  bs <- z[grepl("brown sugar", z)]
  #if no ingredient of interest is found, an empty character will be returned
  #change the empty character to string "missing"
  brown_sugar[i] <-  ifelse(identical(bs, character(0)), "missing",bs)
}
brown_sugar

```

Now that we've seen one way to isolate an ingredient, we can turn it into a function that takes as input an ingredient string (e.g. "egg") and the list-of-ingredients lists (the list *ingredients* from before; it's actually an N-length list of many character vectors). This function returns an N-length character vector of the ingredient and quantity that matches the input ingredient. 

```{r}

#takes as input list of N dataframes of ingredients, and a single ingredient to search for; 
#returns N-length vector of (first match) given ingredient to search for
single_ingredient <- function(ingredientlist_list, single_ingredient){
  single_ingredient %<>% as.character()
  ingred <- vector(mode="character", length=length(ingredientlist_list))
  for (i in 1:length(ingredientlist_list)){
    z <- ingredientlist_list[[i]]    #pick the recipe character list to search over
    bs <- z[grepl(single_ingredient, z)]  #isolate the ingredient 
    ingred[i] <-  ifelse(identical(bs, character(0)), "missing",bs)
  }
  return(ingred)
}

egg_test <- single_ingredient(ingredients, "egg")
knitr::kable(egg_test[1:12])

```

We can then use this function to create a data frame with multiple ingredients:

```{r}

#create data frame of 
z <- data.frame(brown_sugar = single_ingredient(ingredients, "brown sugar"),
                white_sugar = single_ingredient(ingredients, "white sugar"),
                egg = single_ingredient(ingredients, "egg"),
                flour = single_ingredient(ingredients, "flour"))
#turn into characters
z[] <- lapply(z, as.character)

#view view few rows
knitr::kable(z[1:5,])

```

The unicode fractions will cause trouble, so we can write a function to replace them all

```{r}

clean_unicode_fractions <- function(x){
  x <- as.character(x)
  x <- gsub("⅛", ".125", x)
  x <- gsub("¼", ".25", x)
  x <- gsub("½", ".5", x)
  x <- gsub("¾", ".75", x)
  x <- gsub("<U+2009>", "", x) #remove small unicode space
  x <- gsub("\u2009", "", x) #remove small unicode space
  x <- gsub("\u2153", ".333", x) #remove unicode 1/3
  x <- gsub("\u2154", ".667", x) #remove unicode 2/3
  x <- ifelse(startsWith(x, "."), paste0("0",x),x)  #if fraction starts with point, add leading 0
  return(x)
}

```

We use this new unicode-function to clean and extract values from the two sugars ingredients columns:

```{r}

#Create columns that store value-only for amounts
z$w_sugar_amt <- z$white_sugar %>% clean_unicode_fractions() %>% 
  str_extract("\\d+\\.*\\d*") %>% as.numeric()
z$b_sugar_amt <- z$brown_sugar %>% clean_unicode_fractions() %>% 
  str_extract("\\d+\\.*\\d*") %>% as.numeric()


#view view few rows
knitr::kable(z[1:15,])

```

We can see a potential problem: the function does not care about units, so "2 tablespoons of sugar" and "2 cups of sugar" will both return a 2. An easy way to get around this is to limit our recipes to only those who use "cup" as the measurement in both sugars columns. Using *dplyr*s filter as below will also only keep recipes that have both brown and white sugars:

```{r}

#Keep complete recipes:
zz <- filter(z, grepl("cup", brown_sugar)& grepl("cup", white_sugar))

```

We can do a basic scatter plot using ggplot of the amounts of white-versus-brown sugar in chocolate chip cookie recipes. Note that overplotting (points on top of points) is a problem with this data set, so we can use the "position_jitter" option and set a transparency alpha so that we can get an idea of how many recipes use the most common ratios (for instance, the classic 0.75-0.75 cup white-brown sugar ratio of the original Nestle Tollhouse recipe).

```{r}

#Plots
ggplot(zz, aes(x=b_sugar_amt, y=w_sugar_amt)) + 
  geom_point(position = position_jitter(h= 0.025, w = 0.025), alpha = 0.3, size = 2) + 
  geom_abline(slope = 1) + 
  ggtitle("Brown vs. White Sugar in Choc. Chip Cookie Recipes") +
  xlab("Brown Sugar (cups)") +
  ylab("White Sugar (cups)")+ 
  coord_fixed(ratio = 1, xlim = NULL, ylim = NULL, expand = TRUE, clip = "on")

```

We can also create a ratio of the two sugars and graph that:

```{r}

zz$sugar_ratio <- zz$b_sugar_amt / zz$w_sugar_amt
hist(zz$sugar_ratio, n = 50, 
     main ="Sugar Ratio", 
     xlab = "Ratio of white:brown sugar")

```


```{r, eval =  F, echo = F}

fig <- plot_ly(type = 'scatter', mode = 'markers') 
fig <- fig %>%
  add_trace(
    x = z$b_sugar_amt, 
    y = z$w_sugar_amt, 
    marker = list(color='green'),
    hoverinfo = z$brown_sugar,
    showlegend = F
  ) 
fig <- fig %>%
  layout(
    title = "Sugar in Cookie recipes",
    xaxis = list(zeroline = F),
    yaxis = list(hoverformat = '.2f'))

fig

```

